---
title: "PCV antibodies"
author: "Dan Weinberger"
date: "9/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(reshape2)
library(shiny)
library(ggplot2)
```

```{r}
pcv7sts <- c('4','6B','9V','14','18C','19F','23F')
d1a <- read_excel("./Data/IgGGMCs2.xlsx")
d1 <- d1a[d1a$Trial!='US2',]
names(d1)[1] <- 'vax'
d1$timepoint_dose_weeks <- NULL
d1.m <- melt(d1, id.vars=c('vax','Dose','Trial'))
d1.m$variable <- gsub('serotype', '', d1.m$variable)
d1.m <- d1.m[d1.m$variable %in% pcv7sts ,]
d1.c <- acast(d1.m, vax~Dose~Trial+variable)

boost <- as.data.frame(t(d1.c[,'Booster',]))

d2 <- dcast(d1.m, vax+Dose+Trial+variable~.)
names(d2) <- c('vax','Dose','Trial','st','gmc')

```

```{r}
plot(boost$PCV7, boost$PCV13, bty='l')
abline(a=0, b=1)
```

## From Voysey CID
Correlate of protection carriage
ST4: 1.16
ST6B 0.5
ST9V 1.31
ST14 2.48
ST19F 2.54
ST23F 0.63
```{r}

post.prim <- as.data.frame(t(d1.c[,'3rddoseP',]))
post.prim <- post.prim[!is.na(post.prim$PCV7) & !is.na(post.prim$PCV13),]

#prim2 <- as.data.frame(t(d1.c[,'2nddoseP',]))
#prim2 <- prim2[!is.na(prim2$PCV7) & !is.na(prim2$PCV13),]

pre.boost <- as.data.frame(t(d1.c[,'preBooster',]))
pre.boost <- pre.boost[!is.na(pre.boost$PCV7) & !is.na(pre.boost$PCV13),]

```



#PCV7 13 ratio for post-primary, pre/post booster
```{r, fig.width=3, fig.height=8}
par(mfrow=c(3,1), mar=c(4,4,1,1))
all.ds <- list( 'post.prim'=post.prim,'pre.boost'=pre.boost, 'boost'=boost)
ds.labels= c('Post primary','Pre-Booster','Post-booster')

for(i in 1:length(all.ds)){
  ds <- all.ds[[i]]
ratio <- ds$PCV13/boost$PCV7
study <- gsub( "_.*$", "", rownames(ds) )
st <- gsub( ".*_", "", rownames(ds) )

plot.df <- cbind.data.frame(ratio, study, st)

plot(as.numeric(as.factor(plot.df$st)), plot.df$ratio, xaxt='n', bty='l', ylab='Ratio', main=ds.labels[i] , ylim=c(0,1.5))
axis (side=1, at=c(1:length(unique(plot.df$st))), labels=unique(plot.df$st) )
abline(h=1, lty=2)
}

```
Quantify differences in a regression. Note this doesn't account for differing uncertainty between studies.

```{r}
names(d1a)[1] <-'vax'

d1a.m <- melt(d1a,id.vars = c('vax','Dose','Trial', 'timepoint_dose_weeks'))
d1a.m$value[d1a.m$value==0] <- 0.01 
d1a.m$log.ab <- log(d1a.m$value)
d1a.m$Dose <- as.factor(d1a.m$Dose)
d1a.m$timepoint_dose_weeks <- as.factor(d1a.m$timepoint_dose_weeks)
d1a.m$vax <- as.factor(d1a.m$vax)
d1a.m$vax <- relevel(d1a.m$vax, 'PCV7')

d1a.m$st <- as.factor(d1a.m$variable)

d1a.m.spl <- split(d1a.m,d1a.m$Dose)
mod1 <- lm(log.ab ~  timepoint_dose_weeks + vax  +st + vax*st, data=d1a.m.spl[["3rddoseP"]]) 
summary(mod1)


```

```{r}
d1.m.boost <- d1.m[d1.m$Dose=='Booster',]
d1.c.alt <- dcast(d1.m.boost, vax~Trial+variable)

d1.c.alt$vax.d <- 0
d1.c.alt$vax.d[d1.c.alt$vax=='PCV13'] <- 1
d1.c.alt <- d1.c.alt[c(2,1),]

matplot(log(d1.c.alt[,-1]), type='l', col=rgb(0,0,0,alpha=0.1), lty=1)

```

## Siber
https://www.sciencedirect.com/science/article/pii/S0264410X07001569?via=ihub

This is if we want to derive the correlate of protection based on immunogenicity and efficacy data; or if we want to see the proportion of population over correlate of protection
```{r}

#Northern California pooled serotypes; serotype 19F

N.trial=c(190,189)
ab <- list('vax'=c(1.44, 1.26,1.65),
'unvax'=c(0.11, 0.09, 0.13))

ab <- lapply(ab, function(x) x*1)

# ab <- list('vax'=c(2.54,2.13, 3.02),
# 'unvax'=c(0.03, 0.02, 0.03)) 

correlates.func <- function(ab1, N.people, efficacy=0.846, corr.prot=0.35){
    ab.se <- lapply(ab1, function(x){
      se.est.l <- (log10(x[1])-log10(x[2])) /1.96
      se.est.u <- (log10(x[3])-log10(x[1]))/1.96
      log10.se.est <- mean(c(se.est.u,se.est.l))
      log10.sd.est <- log10.se.est*sqrt(N.people[1])
      
      return(c('mean'=x[1], 'log.mean'=log10(x[1]),'log.se'=log10.se.est,'log.sd'=log10.sd.est))
    })
    
    
    #plot(df1$conc, df1$prop)
    
    xvals <- 10^(seq(-2,2, by=0.02))
    cum.vax <- 1 - pnorm(log10(xvals), ab.se$vax['log.mean'], ab.se$vax['log.sd'])
    cum.unvax <- 1 - pnorm(log10(xvals), ab.se$unvax['log.mean'], ab.se$unvax['log.sd'])
    
    pct.protected <-1 - pnorm(log10(corr.prot), ab.se$vax['log.mean'], ab.se$vax['log.sd'])

        dens1 <- dnorm(log10(xvals), ab.se$vax['log.mean'], ab.se$vax['log.sd'])
        dens1 <- dens1/sum(dens1)
        
                dens2 <- dnorm(log10(xvals), ab.se$vax['log.mean']+log10(0.7), ab.se$vax['log.sd'])
        dens2 <- dens2/sum(dens2)

        
    #Compare to Fig 2 in Black, PIDJ
    plot(log10(xvals), cum.vax, type='l', col='red', xaxt='n',xlab='[Ab]', ylab='Proportion', bty='l')
    points(log10(xvals), cum.unvax, type='l', col='blue')
    abline(v=log10(c(0.15,0.5,1.0)), lty=3)
    axis(side=1, at=log10(c(0.01,0.1,1,10,100)), c(0.01,0.1,1,10,100))
    legend(x=0.7, y=1, legend=c('Vaccine','Control'),lty=1, col=c('red','blue'), bty='n')
    
#Histogram   
    par(mfrow=c(2,1), mar=c(2,2,1,1))
    plot(log10(xvals), dens1,xaxt='n',xlab='[Ab]', 
         ylab='Proportion', bty='l', type='l', yaxs="i")
        axis(side=1, at=log10(c(0.01,0.1,1,10,100)),
             c(0.01,0.1,1,10,100))
     # abline(v=log10(0.35), lty=2, col=rgb(0,0,0,alpha=0.2))  
      polygon( log10(c(0.3,0.3, 0.7,0.7)), 
               c( 0,1, 1,0) ,
               col=rgb(0,0,0,alpha=0.3), 
               border=NA )
      shadex <- log10(xvals)[log10(xvals)<log10(0.35)]
      shadex2 <- log10(xvals)[log10(xvals)>=log10(0.35)]

            polygon(c(shadex, rev(shadex)), c(dens1[1:length(shadex)], 
                                        rep(0, length(shadex)) ),
                    col=rgb(1,0,0,alpha=0.5), border=NA  )
                        polygon(c(shadex2, rev(shadex2)),
                                               c(dens1[(length(shadex)+1):(length(dens1))], 
                                        rep(0, length(shadex2)) ),
                    col=rgb(0,0,1,alpha=0.5), border=NA  )
                        
      #add 2nd
    plot(log10(xvals), dens2,xaxt='n',xlab='[Ab]', 
         ylab='Proportion', bty='l', type='l', yaxs="i")
        axis(side=1, at=log10(c(0.01,0.1,1,10,100)),
             c(0.01,0.1,1,10,100))
      abline(v=log10(0.35), lty=2, col=rgb(0,0,0,alpha=0.2))  
  polygon(c(shadex, rev(shadex)), c(dens2[1:length(shadex)], 
                                        rep(0, length(shadex)) ),
                    col=rgb(1,0,0,alpha=0.5), border=NA  )
                        polygon(c(shadex2, rev(shadex2)),
                                               c(dens2[(length(shadex)+1):(length(dens2))], 
                                        rep(0, length(shadex2)) ),
                    col=rgb(0,0,1,alpha=0.5), border=NA  )
     
      #INSTEAD DO viOILIN
  #     p.ds1 <- cbind.data.frame('vax'='PCV13','val'=rnorm(10000,ab.se$vax['log.mean'], ab.se$vax['log.sd']))
  #     p.ds2 <- cbind.data.frame('vax'='PCV20','val'=rnorm(10000,ab.se$vax['log.mean']+log10(0.5), ab.se$vax['log.sd']))
  #     p.ds.c <- rbind.data.frame(p.ds1, p.ds2)
  #     p.ds.c$vax <- as.factor(p.ds.c$vax)
  #     p.ds.c.top <- p.ds.c[p.ds.c$val>log10(0.35),]
  #     p.ds.c.bottom <- p.ds.c[p.ds.c$val<=log10(0.35),]
  #     
  #     p <- ggplot(p.ds.c.top, aes(x=vax, y=val, fill=vax)) + 
  # geom_violin(trim=T, adjust=2)+
  # geom_violin(data=p.ds.c.bottom,trim=T, adjust=2)+
  #               theme_classic()
  #      
    
    #or invert
    plot(log10(xvals), 1-cum.vax, type='l', col='red')
    points(log10(xvals), 1-cum.unvax, type='l', col='blue')
    
    
    ratio1 <- 1-(1-cum.vax) / (1-cum.unvax)
    plot(log10(xvals), ratio1)
    
    ve <- efficacy
    plot(log10(xvals),abs(ratio1-ve), ylim=c(-2,2))
    xvals[which(min(abs(ratio1-ve), na.rm=T )==abs(ratio1-ve))]

    return(pct.protected)
    
    
}

d1 <- correlates.func(ab1=ab, N.people = N.trial[1],corr.prot=0.35)
d1

```

```{r}
#Serotype 1 CAPITA
#https://academic.oup.com/cid/article/65/5/787/3920608#96465713
#N=393
#base.titers <- c(10, 9.7,10.6) 
#titer30d <- c(107, 90.9,126) 

titers <- list('unvax'=base.titers, 'vax'=titer30d)
ab.sd <- lapply(titers, function(x){
      sd.est.l <- (log10(x[1])-log10(x[2])) /1.96
      sd.est.u <- (log10(x[3])-log10(x[1]))/1.96
      log10.sd.est <- mean(c(sd.est.u,sd.est.l))
      return(c('mean'=x[1], 'log.mean'=log10(x[1]),'log.sd'=log10.sd.est))
    })

xvals <- 10^(seq(0,4, by=0.02))
    cum.vax <- 1 - pnorm(log10(xvals), ab.sd$vax['log.mean'], ab.sd$vax['log.sd'])
    cum.unvax <- 1 - pnorm(log10(xvals), ab.sd$unvax['log.mean'], ab.sd$unvax['log.sd'])

    plot(log10(xvals), cum.vax, type='l', col='red', xaxt='n')
    points(log10(xvals), cum.unvax, type='l', col='blue')
    abline(v=log10(c(0.15,0.5,1.0)), lty=3)
    axis(side=1, at=log10(c(0.01,0.1,1,10,100)), c(0.01,0.1,1,10,100))
    
    
```


##Sample shiny App
```{r period.slider, echo=FALSE}
#library(shiny)

shinyApp(

  ui = fluidPage(
      sidebarPanel(
    selectInput("vax", "Vaccine:",
               unique(d2$vax), multiple=T, selected=unique(d2$vax)),
    checkboxGroupInput("country", "Country:",
               unique(d2$Trial), selected=unique(d2$Trial)),
    checkboxGroupInput("doses", "Doses:",
               unique(d2$Dose), selected=c("3rddoseP" ,"Booster")),
    checkboxGroupInput("st", "Serotypes:",
                unique(d2$st),  selected=unique(d2$st)),
    selectInput('plot_type', 'Plot type:',
                c('Raw Concentration','Ratio'), selected= 'Raw Concentration') 
    ),
    mainPanel(
      plotOutput("periodPlot")
  )
  ), 

  server = function(input, output) {
    output$periodPlot = renderPlot({
      
      if(input$plot_type=='Raw Concentration'){

        plot.ds <- d2[(d2$vax %in% input$vax & 
                        d2$Dose %in% input$doses & 
                        d2$st %in% input$st &
                        d2$Trial %in% input$country ) ,]
      
        p1 <- ggplot(plot.ds, aes(x=vax, y=log(gmc), group=vax, col=vax) ) +
          geom_point() +
          geom_line(aes(group = Trial),color="grey") +
          theme_classic()+
          ylab('log(GMC)') +
          facet_grid(Dose~st ) +
          theme(panel.spacing = unit(1, "lines"))
        p1
  
  
      
     }else{
        plot.ds <- d2[(d2$vax %in% input$vax & 
                        d2$Dose %in% input$doses & 
                        d2$st %in% input$st &
                        d2$Trial %in% input$country ) ,]
        plot.ds.c <- reshape2::dcast(plot.ds, Dose+Trial+st ~vax, value.var='gmc')
        plot.ds.c <- plot.ds.c[!is.na(plot.ds.c$PCV13) & !is.na(plot.ds.c$PCV7),]
         p2 <- ggplot(plot.ds.c, aes(x=st, y=log(PCV13/PCV7), group=Trial, col=Trial) ) +
          geom_point() +
          theme_classic()+
          ylab('log(GMC)') +
         geom_line(aes(group = Trial),color="grey") +
          geom_hline(yintercept=0) +
          facet_grid(Dose~. ) +
          theme(panel.spacing = unit(1, "lines"))
        p2
        
      }
      
      
    },width = "auto", height = "auto")
  }
)
```






